### 2013-05-30 inst/examples/BUGS/allen.Rmd
allen example updated

```{r}
posterior.mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

@@

```{r common-priors}
stdQ_prior_p <- c(0.0001, 100)
stdR_prior_p <- c(1e-6, .1)
stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
```

@@

```{r allen-output, dependson="allen-traces"}
pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable, subset=.(variable!="deviance")) 
pardist[,2] = exp(pardist[,2]) # transform model parameters back first
pardist[,3] = exp(pardist[,3])
bayes_coef <- apply(pardist,2, posterior.mode) 
bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
allen_means <- sapply(x_grid, f, 0, bayes_pars)
```

@@

```{r ricker_traces, dependson="ricker-mcmc"}
tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(ricker_posteriors) = c("index", "variable", "value", "chain")

@@

```{r ricker_posteriors, dependson=c("ricker-traces", "ricker-priors")}
ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
# plot posterior distributions
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

@@

```{r ricker-output, dependson="ricker-traces"}
ricker_pardist <- acast(ricker_posteriors[2:3], 
                        1:table(ricker_posteriors$variable) ~ variable, 
                        subset=.(variable!="deviance"))

ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
```

@@

```{r myers-output, dependson="myers-traces"}
myers_pardist <- acast(myers_posteriors[2:3], 
                        1:table(myers_posteriors$variable) ~ variable, 
                        subset=.(variable!="deviance")) 
myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
```

@@

# labesl for the colorkey too
model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
colorkey=cbPalette
names(colorkey) = model_names 
```

@@

```{r par-fns}
ricker_f <- function(x,h,p){
  sapply(x, function(x){ 
    x <- pmax(0, x-h) 
    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
  })
}
allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])

```{r Figureb, dependson=c("assemble-models", "par-fns")}
require(MASS)
step_ahead <- function(x, f, p){
  h = 0
  x_predict <- sapply(x, f, h, p)
  n <- length(x_predict) - 1
  y <- c(x[1], x_predict[1:n])
  y
}
step_ahead_posteriors <- function(x){
gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
df_post <- melt(lapply(sample(100), 
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Allen = step_ahead(x, allen_f, pardist[i,]), 
                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
  }), id=c("time", "stock"))
}

@@

```{r allen-opt, dependson="allen-output"}
allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r ricker-opt, dependson="ricker-output"}
ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r myers-opt, dependson="myers-output"}
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r}
df <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable) 
modes <- apply(df, 2, posterior.mode)
allen_deviance <- modes[['deviance']]

df <- acast(ricker_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
modes <- apply(df, 2, posterior.mode)
ricker_deviance <- modes[['deviance']]

df <- acast(myers_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
modes <- apply(df, 2, posterior.mode)
myers_deviance <- modes[['deviance']]

true_deviance <- 2*estf(c(p, sigma_g))
mle_deviance <- 2*estf(c(est$p, est$sigma_g))

>>>>

```{r posterior-mode}
require(modeest)
posterior.mode <- function(x) {
  mlv(x, method="shorth")$M
}

@@

```{r common-priors}
stdQ_prior_p <- c(0.0001, 100)
stdR_prior_p <- c(1e-4, .1)
stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
```

@@

```{r allen-output, dependson="allen-traces"}
A <- allen_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
pardist <- acast(A, index ~ variable)
# pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable)[1] ~ variable) # NOT SURE WHY THIS FAILS

# transform model parameters back first
pardist[,"logr0"] = exp(pardist[,"logr0"]) 
pardist[,"logtheta"] = exp(pardist[,"logtheta"])
colnames(pardist)[colnames(pardist)=="logtheta"] = "theta"
colnames(pardist)[colnames(pardist)=="logr0"] = "r0"

bayes_coef <- apply(pardist,2, posterior.mode) 
bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
allen_f <- function(x,h,p) unname(RickerAllee(x,h, unname(p[c("r0", "K", "theta")])))
allen_means <- sapply(x_grid, f, 0, bayes_pars)

bayes_pars
head(pardist)

@@

```{r ricker-traces, dependson="ricker-mcmc"}
tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(ricker_posteriors) = c("index", "variable", "value", "chain")

@@

```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
# plot posterior distributions
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

@@

```{r ricker-output, dependson="ricker-traces"}
ricker_pardist <- acast(ricker_posteriors[2:3], 
                        1:table(ricker_posteriors$variable) ~ variable)

ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
colnames(ricker_pardist)[colnames(ricker_pardist)=="logr0"] = "r0"
bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
ricker_bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"]))

ricker_f <- function(x,h,p){
  sapply(x, function(x){ 
    x <- pmax(0, x-h) 
    pmax(0, x * exp(p["r0"] * (1 - x / p["K"] )) )
  })
}
ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])

head(ricker_pardist)
ricker_bayes_pars
```

@@

```{r myers-output, dependson="myers-traces"}
myers_pardist <- acast(myers_posteriors[2:3], 
                        1:table(myers_posteriors$variable) ~ variable, 
                        subset=.(variable!="deviance")) 
myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c("r0", "theta", "K")])
head(myers_pardist)
myers_bayes_pars

@@

# labels for the colorkey too
model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
colorkey=cbPalette
names(colorkey) = model_names 
```

@@

```{r Figureb, dependson=c("assemble-models")}
require(MASS)
step_ahead <- function(x, f, p){
  h = 0
  x_predict <- sapply(x, f, h, p)
  n <- length(x_predict) - 1
  y <- c(x[1], x_predict[1:n])
  y
}
step_ahead_posteriors <- function(x){
gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
df_post <- melt(lapply(sample(100), 
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Allen = step_ahead(x, allen_f, pardist[i,]), 
                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
  }), id=c("time", "stock"))
}

@@

```{r allen-opt, dependson="allen-output"}
matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r ricker-opt, dependson="ricker-output"}
matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r myers-opt, dependson="myers-output"}
matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r deviances, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
allen_deviance <- pardist[,'deviance']
ricker_deviance <- ricker_pardist[,'deviance']
myers_deviance <- myers_pardist[,'deviance']
true_deviance <- 2*estf(c(p, sigma_g))
mle_deviance <- 2*estf(c(est$p, est$sigma_g))
